---
title: "Class 05"
author: "Abel"
format: gfm
---

R has lot's of ways to make figures and graphs in particlular. One that come with R out the box is called **"bas" R** - the 'plot()' function.

```{r}
plot(cars) 
```

------------------------------------------------------------------------

A very popular package in this area is callled **ggplot2**

Before I can use any add-on package like this I must install it with the `install.packages("ggplot2")` command/function

Then to use hte package I need to load it with a 'library(ggplot2)' call.

```{r}
library(ggplot2)
ggplot(cars) + 
  aes(x=speed, y=dist) + 
  geom_point()
```

For "simple" plots like this one base R code witll be much shorter than ggplot code.

```{r}
library(ggplot2)
ggplot(cars) + 
  aes(x=speed, y=dist) + 
  geom_point() +
  geom_smooth()
```

Every ggplot has at least three layers

\-**data** (data.frame with the number or stuff you want to plot) -**aes**tthetics (mapping of your data colums to your plot) -**geom**s (there are tones of these, basics are `geom_point()`, `geom_line()`, `geom_col()`)

```{r}
head(mtcars)
```
Make a ggplot of the `mtcars` data set using `mpp` vs `disp` and set the size fo the points to the `hp`
set the color to `am`
```{r}
library(ggplot2)
ggplot(mtcars) + 
  aes(x=mpg, y= disp, size=hp, color= am) + 
  geom_point() 
  
```


```{r}
library(ggrepel)

ggplot(mtcars) + 
  aes(x=mpg, y= disp, col=am, label=rownames(mtcars)) + 
  geom_point() +
  facet_wrap(~am) +
  geom_text_repel()
  
```


```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
nrow(genes)

```
There are `r nrow(genes)` genese is this dataaset. 
```{r}
unique(genes$State)
```
The `table()`function is super useful untility to tell me how many entries of each type there are.
```{r}
table(genes$State) / nrow(genes)
```

The functions `nrow()`, and `table()` are ones I want you to know
```{r}
library(ggplot2)
ggplot(genes) +
aes(x=Condition1, y=Condition2) +
geom_point()
    
```
```{r}
    ggplot(genes) + 
    aes(x=Condition1, y=Condition2, col=State) +
    geom_point()
```

```{r}
ggplot(genes) + 
    aes(x=Condition1, y=Condition2, col=State) +
    geom_point() +
scale_colour_manual( values=c("blue","gray","red") )
```

```{r}
ggplot(genes) + 
    aes(x=Condition1, y=Condition2, col=State) +
    geom_point() +
scale_colour_manual( values=c("blue","gray","red") ) +
labs(title = "Gene Expression Changes Upton Drug Treatment", x = "Control (no drug)", y= "Drug Treatment")
```


> key points 

```{r}
p <-ggplot(mtcars)+
  aes(mpg, disp) +
  geom_point()

ggsave("myplot.pdf")
```


```{r}
p
```

